---
title: "Class 8"
author: "Nasha Luevit"
date: "February 5, 2019"
output: html_document
---

## Generate some example data for clustering
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

## Use **kmeans()**
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

Q. How many points are in each cluster?
   30 points in each cluster
Q. What ‘component’ of your result object details
 - cluster size?
    'size'
 - cluster assignment/membership?
    'cluster'
 - cluster center?
    'centers'
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points

```{r}
km$size
```

```{r}
km$cluster
```
```{r}
plot(x, col = km$cluster)
points(km$centers, col = "blue", pch = 15, cex = 2)
```

## Heirarchical Clustering 

# First we need to calculate point (dis)similarity as the Euclidean distance between observations
```{r}
dist_matrix <- dist(x)
```

# The **hclust()** function returns a hierarchical clustering model
```{r}
hc <- hclust(d = dist_matrix)
```

# the print method is not so useful here
```{r}
hc
```

# Create hierarchical cluster model: hc
```{r}
hc <- hclust(dist(x))
```

# We can plot the results as a dendrogram
```{r}
plot(hc)
```

# We can use k=number of cluster groups desired with **cutree()**
```{r}
plot(hc)
abline(h = 6, col = "red")
cutree(hc, h = 6)
```

## Diferent Linkage Methods
```{r}
d <- dist_matrix

hc.complete <- hclust(d, method = "complete")
plot(hc.complete)
hc.average <- hclust(d, method = "average")
plot(hc.average)
hc.single <- hclust(d, method = "single")
plot(hc.single)
```

# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

# Step 2. Plot the data without clustering
```{r}
plot(x)
```

# Step 3. Generate colors for known clusters (so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

# Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
plot(hc)
abline(h = 2, col = "red")
abline(h = 2.5, col = "blue")
```

```{r}
gp2 <- cutree(hc, k=2)
gp3 <- cutree(hc, k=3)
plot(x, col=gp3)
```

# Q. How does this compare to your known 'col' groups?
#   they are similar

## Principal Component Analysis (PCA)

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata) 
```

# Let's do PCA
```{r}
pca <- prcomp(t(mydata), scale = TRUE)
attributes(pca)
```

# A basic PC1 vs. PC2 2D plot
```{r}
plot(pca$x[,1], pca$x[,2])
```

# Percent variance per PC
```{r}
#Variance per PC
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```

# Plot
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```
#   PC1 accounts for almost all of the variation in the data

## A vector of colors for wt and ko samples
```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 

#Click to identify which sample is which
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata)) 
```

## Hands on Section
```{r}
x <- read.csv("UK_foods.csv")
dim(x)
#view file
View(x)
#fix
rownames(x) <- x[,1]
x <- x[,-1]
dim(x)
```

```{r}
barplot(as.matrix(x), col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

# PCA can help analyze this data in a better way
```{r}
pca <- prcomp(t(x))
summary(pca)
```

## Plot PC1 vs. PC2
```{r}
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2", xlim = c(-270, 500) )
text(pca$x[,1], pca$x[,2], colnames(x), col = c("orange", "red", "blue", "darkgreen"))
```

# Variation
```{r}
v <- round(pca$sdev^2/sum(pca$sdev^2) * 100)
v
```

# Plot variation
```{r}
barplot(v, xlab = "Principal Component", ylab = "Percent Variation")
```

# Examine the "loadings"
```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```






