---
title: "Class 6"
author: "Nasha Luevit"
date: "January 24, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: Reading files again 

Here we are going to try using the **read.table()** function and friends to read some example flat files (text files)

```{r}
plot(1:10)
```

Back to file reading
(option+command+i will start new code)

```{r}
read.table("https://bioboot.github.io/bimm143_W19/class-material/test1.txt", header = TRUE, sep = ",")
```

For this common CSV format we can use **read.csv()**

```{r}
file1 <- "https://bioboot.github.io/bimm143_W19/class-material/test1.txt"
data1 <- read.csv(file1) 
data1
```

Practice reading files

```{r}
file2 <- "https://bioboot.github.io/bimm143_W19/class-material/test2.txt"
data2 <- read.csv(file2, sep = "$")
data2
```

```{r}
file3 <- "https://bioboot.github.io/bimm143_W19/class-material/test3.txt"
data3 <- read.table(file3)
data3
```

## Section 2: R Functions

My first silly function
```{r}
add <- function(x, y=1) {
  #sum the input x and y
  x + y
}
```


Try using the function
```{r}
add(1)
```

```{r}
add(1,100)
```

```{r}
add( c(1,2,3), 4)
```

```{r}
#add(1,2,2)
#(too many arguments -> returns error)
```


```{r}
#add(x=1, y="b")
#(arguments don't match)
```


My 2nd Function
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

Test on a small example that we know the answer to 
```{r}
rescale(1:10)
```

```{r}
rescale( c(1,2,NA,3,10) )
```

```{r}
#rescale( c(1,10,"string") )
#doesn't work
```

modify function
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

test modified function
```{r}
rescale2( c(1,2,NA, 3, 10) )
```

new function
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
   print("please dont sing anymore barry!")
 }
 print("I can see it in ...")
 return(answer)
}
```

test new function
```{r}
myans <- rescale3( c(1:6, NA, 10), plot=TRUE )
```

modify 
```{r}
rescale4 <- function(x, na.rm=TRUE, plot=TRUE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
   print("please don't ever sing again!")
 }
 print("I can see it in ...")
 return(answer)
}
```

test
```{r}
rescale4( c(1:6, NA,10) )
```


## Working with the bio3d package

to install this package, command **install.packages("bio3d")** was used 

using package 

```{r}
library(bio3d)
```

```{r}
# read a PDB file from the database
s1 <- read.pdb("4AKE") # kinase with drug
s1
```

Try out provided code
```{r}
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```



